---
layout: post
title: "How to Import Data into R"
author: "Chuan Tang"
date: "February 25, 2016"
output: html_document
excerpt: A summary about various methods to import data into R.
tags: R
comments: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

Importing data into R is the first step for all data analysis. In addition to using the basic `read.table()` command, this article introduces (1) the `readr` package developed by [Hadley Wickham](http://hadley.nz/), (2) import data from internet to R using package `jsonlite`, and (3) connect and fetch data from online database. This article is a summary of completing the "Importing Data into R" course of [DataCamp](https://www.datacamp.com/courses/importing-data-into-r). 

Firstly, we load all packaged needed for this sessions:

```{r loading}
library(readr)
library(jsonlite)
```

> Importing data into R using `readr` package 

The basic R provides `read.table()` to import data. I am using a dataset for [globle airports dump file](https://datahub.io/dataset/global_airports/resource/82b48517-63ed-47d7-be5f-e6775fc015bd) as example here:

Firstly, we need to download the data for local use:

```{r download data, cache=TRUE}

url <- "https://commondatastorage.googleapis.com/ckannet-storage/2012-07-09T214020/global_airports.csv"

sample_path <- file.path("~", "sample.csv")  # set the file path for download data

download.file(url, sample_path)

```
Once we download the data successfully, we can read the downloaded file directly from the `sample_path`. `readr` provides `read_csv` for `.csv` file and `read_tsv` for `.txt` file:

```{r read table using readr}
library(readr)

sample_data <- read_csv(sample_path)

```

Then, we can check the structure of the `example` data:

```{r, echo=TRUE}
str(sample_data)
```

The biggest advantage of using the `read_csv()` rather than `read.csv()` is that the former is more concise and parsimony: you no longer need to specify the `header` and `sep` keys for most situations. Furthermore, as shown above, the data structure includes `tbl_df`, which is convenient for data analysis using package `data.table`. 

> Import **JSON** data into R

[JSON(JavaScript Object Notation)](http://www.json.org/) is a lightweight data-interchange format. The upside of using *JSON* data is it is easy for human reading and writing. 

Firstly, it is necessary to briefly introduce the Syntax rules for *JSON* data:

#### JSON Data

  + Example:  "firstName":"Chuan"

#### JSON Values

  + Example: {"firstName":"Chuan", "lastName":"Tang"}
  
#### JSON Arrays

  + Example: 
  
  "employees":[
    {"firstName":"Chuan", "lastName":"Tang"},
    {"firstName":"Jason", "lastName":"Smith"},
    {"firstName":"Peter","lastName":"Jones"}
]

More details about *JSON* format and how to use it can be found it [here](http://www.w3schools.com/json/default.asp).

The package used for fetching and converting *JSON* data into R is [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/index.html) developed by Jeroen Ooms. I am using Reservoir Information Feed data in *JSON* format published on [datahub](https://datahub.io/dataset/reservoir-information-feed/resource/44ea13ec-c9a8-4fc5-9480-4b9f4fa59718) as example.


At first, we import the online *JSON* data into R using the `fromJSON()` command :

```{r jason import, eval=FALSE}

data_url <- "http://www.ywonline.co.uk/web/reservoi.nsf/ReservoirsJSON"

json_data <- fromJSON(data_url)  # It is possible to directly put the url here as well

str(json_data)
```

As to convert R data into *JSON* format, `toJSON()` command could be used. Moreover, `prettify()` and `minify()` are two commands to make the *JSON* pretty or as concise as possible:

```{r}

pretty_json <- toJSON(mtcars, pretty = T)  # to make JSON data in a pretty way, the pretty key could be set directly

mini_json <- minify(pretty_json)  # to convert the pretty JSON data into concise way

```

The session info for this test is shown below:
```{r}

sessionInfo()

```














